Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest

### 2024-10-22 paper/paper.qmd
committing text changes to the paper.qmd

We use weakly informative Normal priors for the intercept ($\alpha_i$) and the coefficients ($\beta_i$) and $\gamma_i$), centered around zero with a standard deviation of 2.5. These priors were chosen to reflect the assumption that, before seeing the data, there is no strong reason to favor one candidate over the other across states, but to allow room for the data to adjust predictions as polling percentages differ.

We run the model in R [@citeR] using the `rstanarm` package [@rstanarm]. The model uses default priors from `rstanarm`, and estimates the likelihood of Donald Trump winning each state based on polling averages for both candidates.

@@

Bayesian logistic regression was chosen for its ability to model binary outcomes like win/loss predictions while incorporating prior knowledge and uncertainty, helping mitigate overfitting through regularization with priors. This model strikes a balance between flexibility and interpretability, given that polling percentages are continuous and the outcome is binary. The model assumes polling data accurately reflects voter intentions and that state outcomes are independent of each other, although it does not account for interactions between states or regional voting trends. Potential biases in polling data could also affect predictions. The model was implemented using the 'rstanarm' [@rstanarm] package in R[@citeR], which supports Bayesian inference with default priors for logistic regression.

@@

This process helps reveal whether the model is over-fitting to the training data or truly capturing the underlying patterns that apply more broadly. The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction. Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set. The testing was implemented using the following packages in R[@citeR]: caret[@caret], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr].

@@

The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times[@nyt-siena-methodology].

@@

Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election.

To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@sec-references] [@US-election-survey].

@@

Within each state, we will apply stratified sampling again, dividing the population into 6 groups and selecting 100 random samples from each stratum within those groups, ensuring our sample captures key demographic critical for predicting voter behavior. The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other).

Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth. Voting history will be classified into those who voted in the previous election and non-voters, while party registration will include Democrats, Republicans, and Independents.

>>>>

We use weakly informative Normal priors for the intercept ($\alpha_i$) and the coefficients ($\beta_i$) and $\gamma_i$), centered around zero with a standard deviation of 2.5. These priors were chosen to reflect the assumption that, before seeing the data, there is no strong reason to favor one candidate over the other across states, but to allow room for the data to adjust predictions as polling percentages differ. We run the model in R [@citeR] using the `rstanarm` package [@rstanarm]. The model uses default priors from `rstanarm`, and estimates the likelihood of Donald Trump winning each state based on polling averages for both candidates.

@@

Bayesian logistic regression was chosen for its ability to model binary outcomes like win/loss predictions while incorporating prior knowledge and uncertainty, helping mitigate overfitting through regularization with priors. This model strikes a balance between flexibility and interpretability, given that polling percentages are continuous and the outcome is binary. The model assumes polling data accurately reflects voter intentions and that state outcomes are independent of each other, although it does not account for interactions between states or regional voting trends. Potential biases in polling data could also affect predictions. The model was implemented using the 'rstanarm' [@rstanarm] package in R [@citeR], which supports Bayesian inference with default priors for logistic regression.

@@

This process helps reveal whether the model is over-fitting to the training data or truly capturing the underlying patterns that apply more broadly. The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction. Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set. The testing was implemented using the following packages in R [@citeR]: caret [@caret], rstanarm [@rstanarm], ggplot2 [@ggplot2], dplyr [@dplyr].

@@

The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times [@nyt-siena-methodology].

@@

Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@sec-references] [@US-election-survey].

@@

Within each state, we will apply stratified sampling again, dividing the population into 6 groups and selecting 100 random samples from each stratum within those groups, ensuring our sample captures key demographic critical for predicting voter behavior. The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other). Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth. Voting history will be classified into those who voted in the previous election and non-voters, while party registration will include Democrats, Republicans, and Independents.
